---
title: "Quick Start Guide"
author: "Aastha"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# bootplus Quick Start Guide

Welcome to **bootplus**! This guide will show you how to use the three main capabilities of the package in just a few minutes.

## Installation

```{r, eval=FALSE}
# Install from GitHub (devtools required)
devtools::install_github("aastha-chhabra/Bootplus")
```

```{r setup}
library(bootplus)
library(boot)
```

---

## Part 1: Classical Bootstrap with Visualization

The `boot` package is fantastic for resampling, but `bootplus` adds modern visualization and interpretation tools.

### Example: Bootstrap the mean of `mtcars$mpg`

```{r classical-bootstrap, fig.width=7, fig.height=5}
set.seed(42)

# Perform bootstrap
b <- boot(mtcars$mpg, function(d, i) mean(d[i]), R = 2000)

# Visualize the bootstrap distribution
boot_viz(b)
```

The plot shows:
- **Histogram** of bootstrap replicates
- **Density curve** overlaid
- **Original statistic** (vertical black line)
- **Confidence interval** (shaded region)
- **Bias annotation** (how far bootstrap mean is from original)

### Interpret the results

```{r interpret}
# Get a human-readable summary
boot_interpret(b)
```

This tells you:
- Whether bias is negligible or substantial
- The standard error of your estimate
- A tidy confidence interval
- A narrative conclusion

### Compare confidence interval methods

```{r ci-comparison}
# See which CI method gives the widest/narrowest interval
boot_ci_plot(b)

# Get the numbers
boot_compare_ci(b)
```

Different CI methods (Normal, Basic, Percentile, BCa) can give slightly different results. This visualization helps you choose or understand the tradeoffs.

### Full report

For a comprehensive summary, use `boot_report()`:

```{r report}
# One function that does everything
boot_report(b)
```

---

## Part 2: Bayesian Bootstrap

The Bayesian bootstrap (Rubin, 1981) offers a Bayesian perspective on bootstrap inference. Instead of resampling with replacement, we assign random weights (from a Dirichlet distribution) and compute the statistic on the weighted data.

### Example: Bayesian bootstrap of a weighted mean

```{r bayesian-bootstrap, fig.width=7, fig.height=5}
# Some data
x <- c(2, 4, 5, 8, 10)

# A statistic that takes data and weights
weighted_mean <- function(data, weights) {
  sum(data * weights)
}

# Perform Bayesian bootstrap
bb <- boot_bayes(x, weighted_mean, R = 4000, seed = 42)

# Print summary
print(bb)
```

### Visualize the posterior

```{r bayes-viz}
# Plot the posterior distribution with credible interval
boot_bayes_plot(bb)
```

The shaded region shows the 95% credible interval — the Bayesian equivalent of a confidence interval.

---

## Part 3: Using bootplus with your own statistics

Both visualization and interpretation work with *any* statistic you define. Here are some examples:

### Median

```{r median-example, fig.width=7, fig.height=5}
b_median <- boot(mtcars$mpg, function(d, i) median(d[i]), R = 2000)
boot_viz(b_median, title = "Bootstrap distribution of median MPG")
boot_interpret(b_median)
```

### Correlation

```{r correlation-example, fig.width=7, fig.height=5}
b_cor <- boot(mtcars[, c("mpg", "wt")], 
              function(d, i) cor(d[i, "mpg"], d[i, "wt"]), 
              R = 2000)
boot_viz(b_cor, title = "Bootstrap distribution of correlation")
boot_interpret(b_cor)
```

### Custom statistic: Ratio of means

```{r ratio-example}
ratio_means <- function(d, i) {
  mean(d[i, "mpg"]) / mean(d[i, "wt"])
}
b_ratio <- boot(mtcars[, c("mpg", "wt")], ratio_means, R = 2000)
boot_viz(b_ratio)
```

---

## Key Functions at a Glance

### Visualization
- **`boot_viz()`** — Full diagnostic plot (histogram + density + CI + bias)
- **`boot_density()`** — Streamlined density plot only
- **`boot_ci_plot()`** — Compare CI methods visually
- **`boot_dist_plot()`** — See all replicates as a tile strip

### Interpretation
- **`boot_interpret()`** — Structured text summary
- **`boot_compare_ci()`** — Tidy data frame of CI widths
- **`boot_report()`** — Complete report (interpretation + CI comparison + diagnostics)

### Bayesian Bootstrap
- **`boot_bayes()`** — Perform Bayesian bootstrap
- **`boot_bayes_plot()`** — Visualize posterior with credible interval

---

## Next Steps

- Check `?boot_viz` (or any function) for full parameter details
- Look at the examples in each function's help file
- Visit the [GitHub repository](https://github.com/aastha-chhabra/Bootplus) for issue reports and discussions

## References

- Davison, A.C. & Hinkley, D.V. (1997). *Bootstrap Methods and Their Application*. Cambridge University Press.
- Efron, B. & Tibshirani, R.J. (1993). *An Introduction to the Bootstrap*. Chapman & Hall/CRC.
- Rubin, D.B. (1981). The Bayesian bootstrap. *Annals of Statistics*, 9(1), 130–134.

---

Happy bootstrapping! 🎉
